A Standard for the Scholarly Citation of Archaeological Data

### 2017-09-16 analysis/paper/paper.Rmd
finalised submission to AAP

---
title: "A Standard for the Scholarly Citation of Archaeological Data"
author:
  - "Ben Marwick"
  - "Suzanne E. Pilaar Birch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

structured_data_files_per_file <- 
dc_arch %>% 
    mutate(has_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|pdf", tolower(format)), "has_data", "no")) %>% 
  filter(has_data == "has_data") %>% 
  mutate(format_simple = gsub("application|text|image|\\(|\\)", "", format)) %>% 
  separate(format_simple, 
           sapply(1:50, function(i) paste0('f', i)),
           sep = "/|,|:|;|-|\\.| ") %>% 
  select(datacentre, dataset_id, f1:f50) %>% 
  gather(key = variable, 
         value = value, 
         -datacentre,
         -dataset_id,
         na.rm = TRUE) %>% 
  select(-variable) %>% 
  filter(value != "") %>% 
  mutate(format = str_trim(tolower(value))) %>% 
  filter(grepl("[a-z]+", format)) %>% 
  select(-value) %>% 
  mutate(format = gsub('ms', "", format)) %>% 
  mutate(format = if_else(format %in% c("xls", "xlsx", "excel"),
                          "xls or xlsx", 
                          format)) %>% 
  filter(format != "") %>% 
  distinct(datacentre, dataset_id, format)

@@

top_file_formats <- 
  structured_data_files_per_file %>% 
  group_by(format) %>% 
  tally(sort = TRUE) %>% 
  slice(1:20)

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011). List of turstowrthy repositories are maintained by the editors of the journal _Scientific Data_ (n.d) and <http://www.re3data.org>, a global registry of research data repositories (Pampel et al. 2013). There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.). Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

Data sharing is a relatively new behaviour for most archaeologists, so it is important to clarify what does not count as data sharing. Making data available on a web site that is not a trustworthy, persistent-identifier-issuing repository is not data sharing. For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any tim, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible. Sharing data via supplementary files attached to a journal article is problematic because the are inaccessible to non-subscribers for paywalled journals, and because publishers often alter the data file formats and and file names during the production process. For example, machine-readable format such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files. The modification of supplementary file names during the article production process can break scrips that attempt to read files by referring to file names as they were prior to the publisher handling the files. Finally, authors who write in their publications that 'the data are available by request' are not sharing data. Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.  A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data reposity, and use persistent links to the files in reports and manuscripts.

@@

The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats, except that the publisher may not be a traditional book or journal publisher, but a dedicated data repository organisation. 'Creator' is equivalent to the author or primary collector and arranger of the data (e.g. Open Context or tDAR). The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms. The identifier is a special type of persistent internet address. The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes. There are several varieties of these in use, with DOIs (digital object identifiers) currently the most common form. When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.

@@

- the proxy, usually https://doi.org but sometimes also https://hdl.handle.net/ or just doi:
- the prefix, usually a string of numbers that is unique to the organisation publishing the DOI. For example, tDAR prefix is 10.6067
- the suffix, a random-looking string of any combination of characters and numbers that uniquely identified the file or bundle of files. This is also known as a local identifier (local ID) or an 'accession', an identifier which, by itself, is only guaranteed to be locally unique within the database or source (McMurry et al. 2017)

A DOI alone is not suffcient for citing a dataset because, like many things onthe internet, the files can change over time. For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL the can be shared with reviewers but it not publicly accessible or discoverable by a search engine). Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyses requested by the reviewers. We still have the same DOI in our manuscript for this data set, but the files have changed because of the revisions we did in response to the peer reviewers' comments. Most data repositories will keep each version of the files that you upload, so in this scenario there will be at least two versions of the files in our data repository. Even if we do not follow this exact sequence of updating the files, we may still change them in the future, after our paper has been published. Because of this potential to change the files stored at the DOI, it is necessary to identify specific file versions in our citation of the data. A simple method to record the version is to include the date that the data files were accessed, as when citing any other resource found on the internet. Some repositories also provide DOIs that have a version number suffix like '.v2', '.2', '/2'. However this is not formally a functionality of the DOI system, so including the date accessed is a more robust and general method of indicating the version.

@@

“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://dx.doi.org/10.17605/OSF.IO/RTZTH (Marwick 2017).

@@

Marwick, B. (2017). Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/RTZTH, Open Science Framework, Accessed 7 Sept 2017

@@

One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use. In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible hertiage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007). Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g. graduate degrees, employment, publications, grants, promotions). On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial and technical barriers to accessing that data (e.g. subscription fees to journals). On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits. Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a committement to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.

In short, there is more to the ethics of data sharing than avoiding risk of damage. Redacting the locations of archaeological sites to prevent looting as an obvious case of preparing data for sharing to reduce the risk of damage to the sites. However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012). It is crucial for archaeologists to discuss data sharing plans during negotiatations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation. This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results. The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the ourcome of negotiations about data sharing. If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.

@@

The status of research data is slowly changing in many fields from a privately-held commodity traded for co-authorship and other status tokens to part of the public commons, accessible to anyone with an internet connection. This change in status has been fastest in 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003). However, in 'smaller' sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated. Our pilot studies, while small and limited in many ways, support this with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations. In exmanining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets. Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV. We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to fees charged by a respository.

Among the many challenges of increased the rate of data sharing in archaeology other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required. The challenge of funding can be addressed by professional societies, such as the Society of American Archaeology, communicating to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing and requesting a chance in policy. The challenge of requiring data sharing can also be addressed by funding bodies, but may be most effectively tackled by journal editors and editorial boards, for example by adopting the Transparency and Openness Protocols of the Centre for Open Science (Nosek et al. 2015). Changes in journal-specific submission guidelines can result in desirable changes in authors’ research practices (Giofrè et al 2017). For example, if archaeology journals require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published. As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysied in the manuscript.

The challenge of data sharing as an unrewarded behaviour was the primary motivation for this article. By specifying a standard for the scholarly citation of archaeological data we provide a method for giving credit to the authors or compilers of the data. This rewards those authors with visibility and recognition for their effort with citations, a currency that is valued broadly in the discipline. The data citation standard outlined above is easy for humans and search engines to understand, and is highly functional, with a full internet address to the data contained in the citation. The standard can accomodate the addition of additional information required by different citation styles and publishers' requirements. The data citation standard is not dependent on any specific choices of software, archive, data producer, publisher, or author and is easy to adapt into existing writing and research workflows.

@@

@@

>>>>

---
title: "A Standard for the Scholarly Citation of Archaeological Data"
author:
  - "Ben Marwick, Department of Anthropology, University of Washington, Seatll, 98115, WA, USA, bmarwick@uw.edu, corresponding author"
  - "Suzanne E. Pilaar Birch, Anthropology and Geography Department, University of Georgia,  Athens, Georgia,  30602, USA, sepbirch@uga.edu"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
    fig_caption: yes
    reference_docx: ../templates/template.docx
abstract: |
  "How do archaeologists share their research data, if at all? We review what data are, according to current influential definitions, and previous work on the benefits, costs and norms of data sharing in the sciences broadly. To understand data sharing in archaeology, we present the results of three pilot studies: requests for data by email; review of data availability in published articles, and analysis of archaeological datasets deposited in repositories. We find that archaeologists are often willing to share, but discipline-wide sharing is patchy and ad hoc. Legislation and mandates are effective at increasing data-sharing, but editorial policies at journals lack adequate enforcement. Although most of data available at repositories are licensed to enable flexible reuse, only a small proportion of the data are stored in structured formats for easy reuse. We present some suggestions for improving the state of date sharing in archaeology, among these is a standard for citing data sets to ensure that researchers making their data publicly available receive appropriate credit."
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

structured_data_files_per_file <- 
dc_arch %>% 
    mutate(has_data = ifelse(grepl("access|mdb|dbf|dbase|accdb|dbase|arcgis|shape|shp|mapinfo|mid|mif|surfer|xls|excel|csv|spreadsheet|spreadsheet|pdf", tolower(format)), "has_data", "no")) %>% 
  filter(has_data == "has_data") %>% 
  mutate(format_simple = gsub("application|text|image|\\(|\\)| en ", "", format)) %>% 
  separate(format_simple, 
           sapply(1:50, function(i) paste0('f', i)),
           sep = "/|,|:|;|-|\\.| ") %>% 
  select(datacentre, dataset_id, f1:f50) %>% 
  gather(key = variable, 
         value = value, 
         -datacentre,
         -dataset_id,
         na.rm = TRUE) %>% 
  select(-variable) %>% 
  filter(value != "") %>% 
  mutate(format = str_trim(tolower(value))) %>% 
  filter(grepl("[a-z]+", format)) %>% 
  select(-value) %>% 
  mutate(format = gsub('ms', "", format)) %>% 
  mutate(format = if_else(format %in% c("xls", "xlsx", "excel"),
                          "xls or xlsx", 
                          format)) %>% 
  filter(format != "") %>% 
  distinct(datacentre, dataset_id, format)

@@

top_file_formats <- 
  structured_data_files_per_file %>% 
  group_by(format) %>% 
  tally(sort = TRUE) %>% 
  slice(1:20) %>% 
  filter(format != "p")

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011). List of trustworthy repositories are maintained by the editors of the journal _Scientific Data_ (n.d) and <http://www.re3data.org>, a global registry of research data repositories (Pampel et al. 2013). There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.). Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing practises fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

Data sharing is a relatively new behaviour for most archaeologists, so it is important to clarify what does not count as data sharing. Making data available on a web site that is not a trustworthy, persistent-identifier-issuing repository is not data sharing. For example, data files hosted on a personal faculty webpage or the web server of a research project is problematic because these files could disappear at any time, the data can change in untracked ways, and published links to the files are broken when these websites are reorganised, making the files inaccessible. Sharing data via supplementary files attached to a journal article is problematic because they are inaccessible to non-subscribers for paywalled journals, and because publishers often alter the data file formats and file names during the production process. For example, machine-readable format such as CSV are sometimes presented as PDF or Microsoft Word files in supplementary files. The modification of supplementary file names during the article production process can break scripts that attempt to read files by referring to file names as they were prior to the publisher handling the files. Finally, authors who write in their publications that 'the data are available by request' are not sharing data. Not only are the data not actually available, but there is no way to enforce that the authors will actually honour any requests for their data.  A simple strategy for avoiding these pseudo-sharing behaviours is to deposit data files in a trustworthy data repository, and use persistent links to the files in reports and manuscripts.

@@

The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats, except that the publisher may not be a traditional book or journal publisher, but a dedicated data repository organisation. 'Creator' is equivalent to the author or primary collector and arranger of the data (e.g. Open Context or tDAR). The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms. The identifier is a special type of persistent internet address. The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher, but not necessarily in a human-readable format), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes. There are several varieties of these in use, with DOIs (digital object identifiers) currently the most common form. When a data publisher changes the internet address of a data set, they are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.

@@

- the proxy, usually https://doi.org but sometimes also https://hdl.handle.net/ or just doi:
- the prefix, usually a string of numbers that is unique to the organisation publishing the DOI. For example, tDAR prefix is 10.6067
- the suffix, a random-looking string of any combination of characters and numbers that uniquely identified the file or bundle of files, this is also known as a local identifier (local ID) or an 'accession', an identifier which, by itself, is only guaranteed to be locally unique within the database or source (McMurry et al. 2017)

A DOI alone is not sufficient for citing a dataset because, like many things onthe internet, the files can change over time. For example we might upload data files to the repository when we submit a paper for peer review, then get a DOI and cite that in our manuscript so the peer reviewers can see them (some services provide a private URL that can be shared with reviewers but is not publicly accessible or discoverable by a search engine). Then when we resubmit our manuscript after peer review, we might update the repository to include additional data files relating to additional analyses requested by the reviewers. We still have the same DOI in our manuscript for this data set, but the files have changed because of the revisions we did in response to the peer reviewers' comments. Most data repositories will keep each version of the files that you upload, so in this scenario there will be at least two versions of the files in our data repository. Even if we do not follow this exact sequence of updating the files, we may still change them in the future, after our paper has been published. Because of this potential to change the files stored at the DOI, it is necessary to identify specific file versions in our citation of the data. A simple method to record the version is to include the date that the data files were accessed, as when citing any other resource found on the internet. Some repositories also provide DOIs that have a version number suffix like '.v2', '.2', '/2'. However this is not formally a functionality of the DOI system, so including the date accessed is a more robust and general method of indicating the version. Technological solutions to issues of transience, permanence and granularity in data citation are still in development.

@@

“The data underpinning the analysis reported in this paper were deposited in the Open Science Framework Repository at http://dx.doi.org/10.17605/OSF.IO/32A87 (Marwick 2017).

@@

Marwick, B. (2017). Dataset and R code for "Movement of lithics by trampling: An experiment in the Madjedbebe sediments", http://doi.org/10.17605/OSF.IO/32A87, Open Science Framework, Accessed 7 Sept 2017.

@@

One of these differences is that Western researchers often have individual autonomy in sharing data, and once data are shared, it is free for all to use. In contrast, within indigenous communities, data are part of the group's identity and property, akin to tangible heritage (Nicholas 2014), with certain members entrusted to keep that data on behalf of the group, and sometimes it may be inappropriate for these people to share the data with other members of the group, or beyond the group (Tsosie 2007). Differences between Western and indigenous concepts or data are not only cultural - substantial inequalities exist in access to resources for converting archaeological data into economic and political benefit (e.g. graduate degrees, employment, publications, grants, promotions). On one hand, making archaeological data openly available benefits indigenous communities, because it removes some of the financial and technical barriers to accessing that data (e.g. subscription fees to journals). On the other hand, this may be little more than a useless courtesy (Hymes 1972), because simply making data openly available does not give indigenous people access to the contexts where the data can be used to generate economic and political benefits. Thus, open data has different values for Western and indigenous communities, and if archaeologists want to make a compelling case to their indigenous partners for data sharing, they should also make a commitment to reducing some of the inequalities in making use of the data, for example, by providing some training in data analysis, or adapting the data into a context that is valued by the indigenous community.

In short, there is more to the ethics of data sharing than avoiding risk of damage. Redacting the locations of archaeological sites to prevent looting is an obvious case of preparing data for sharing to reduce the risk of damage to the sites (and in some juridictions, mandated by law). However, because of cultural, economic and political differences between Western archaeologists and the communities they work with, an archaeologist cannot consider their own risk assessment to be comprehensive and exhaustive (Nicholas and Wylie 2012). It is crucial for archaeologists to discuss data sharing plans during negotiations with representatives authorized by the indigenous peoples whose cultural heritage is the subject of investigation. This negotiation process is important because only indigenous peoples themselves can identify potential adverse outcomes to data sharing, and they can do this only if they understand the proposed research and anticipated results. The concrete outcome of this process is often a Memorandum of Understanding between the indigenous community and the archaeologist's employer that documents the outcome of negotiations about data sharing. If the archaeologist's indigenous partners have reasons not to share data in public, then the archaeologist should respect that position, and note in publication that the data are restricted at the request of the indigenous collaborators.

@@

The status of research data is slowly changing in many fields from a gift-culture where data are privately-held commodities traded for co-authorship, invitations, and other status tokens (Hagstrom 1982; Wallis et al. 2013), to part of the public commons, accessible to anyone with an internet connection. This change has been fastest in data-rich 'big science' fields, such as astronomy, physics, and oceanography, that use expensive instruments that generate data shared with large teams of collaborators to produce new discoveries (Reichman and Uhlir 2003). However, in 'smaller' data-poor sciences such as archaeology, where collaborative groups are small, data sharing is still mostly ad hoc and disaggregated. Our pilot studies, while small and limited in many ways, support this statement with an overall sharing rate of `r sharing_rate`% in responses to private requests for data, and `r data_are_avail_perc`% of sampled journal articles with openly available data, in a wide variety of formats and locations. In examining archaeological data deposited in repositories, we found that most of the contents are PDF files of reports, and only a very small fraction of the data files are structured data such as spreadsheets. Archaeologists still have a long way to go in making their data easier to reuse by sharing data in plain text structured data formats such as CSV. We found that legislation requiring data sharing resulted in a substantial increase in the rate of data deposition, and that archaeologists are sensitive to fees charged by a repository.

Among the many challenges of increasing the rate of data sharing in archaeology and other 'small' sciences is that data sharing is unfunded, unrewarded and is only rarely required. The challenge of funding can be addressed by professional societies, such as the Society of American Archaeology, communicating to funding bodies, such as the National Science Foundation, about the importance of allocating funds specifically for data sharing and requesting a chance in policy. The challenge of requiring data sharing can also be addressed by funding bodies, but may be most effectively tackled by journal editors and editorial boards, for example by adopting the Transparency and Openness Protocols of the Centre for Open Science (Nosek et al. 2015). Changes in journal-specific submission guidelines can result in desirable changes in authors’ research practices (Giofrè et al 2017). For example, if archaeology journals adopt incentives such as badging (Kidwell et al. 2016), and require as a condition of publication that submitted articles include in the text a properly formatted citation to the data analysed in the article (or an explanation why the data are not available), authors would quickly adapt to this new requirement to ensure their article will be published. As peer reviewers, we can also communicate to editors that we require the manuscripts that we review to include citations of datasets analysied in the manuscript.

The challenge of data sharing as an unrewarded behaviour was the primary motivation for this article. By specifying a standard for the scholarly citation of archaeological data we provide a method for giving credit to the authors or compilers of the data. This rewards those authors with visibility and recognition for their effort with citations, a currency that is valued broadly in the discipline. The data citation standard outlined above is easy for humans and search engines to understand, and is highly functional, with a full internet address to the data contained in the citation. The standard can accommodate the addition of additional information required by different citation styles and publishers' requirements. The data citation standard is not dependent on any specific choices of software, archive, data producer, publisher, or author and is easy to adapt into existing writing and research workflows.

@@

Hagstrom, W O (1982). Gift Giving as an Organising Principle in Science In: Barnes, B and Edge, D eds.  Science in Context: Readings in the Sociology of Science. Milton Keynes, United Kingdom: The Open University Press, pp. 21–34.

@@

Sawyer, S. (2008). Data wealth, data poverty, science and cyberinfrastructure 1. Prometheus, 26(4), 355-371.
